\chapter{Related Work}\label{ch:related-work}

Numerous studies have focused on various topics in the fast-developing field of wireless communications and networking, from handover prediction to user mobility and network traffic prediction.
This chapter gives a summary of the important research that is related to this thesis.



\section{Handover Prediction}

Montavont et al.~\cite{handover-assisted-by-gps} propose a handover decision algorithm based on the \ac{gps} location of the mobile device.
Unfortunately, in large-scale and dense \ac{wifi} environments, \ac{gps} may not be available or not accurate enough for indoor trajectories.
Khan et al.~\cite{MLBasedHandoverPrediction2022} address the problem of handover prediction and \ac{ap} selection in dense \ac{wifi} networks with \ac{sdn}.
Their \ac{ap} selection predictions outperform the current approaches of strongest received signal first by 9.2\,\% and least-loaded \ac{ap} first by 8\,\%.
Khan et al. focus on using \ac{ml} for throughput estimation of the network and accordingly choose the best \ac{ap} to roam to; they do not consider the trajectory of the mobile device in the \ac{ap} selection or use \ac{ml} for the \ac{ap} selection process directly.

\section{User Movement Prediction}

Bakirtzis et al.~\cite{multivariate-lstm-indoor-outdoor} treat their indoor-outdoor detection problem as a multivariate time-series classification.
They use \ac{ml} containing \ac{lstm} for their prediction and demonstrate that a multivariate time-series classification approach can be used to monitor a user's environment.
To predict user movement, Bourjandi et al.~\cite{bourjandiPredictingUserMovement2022} use a mix of multiple \ac{ml} models, where \ac{lstm} is used to learn long-term dependencies.
Prasad et al.~\cite{hmm-movement-prediction} propose a \ac{hmm} to predict the next possible location.
They use real-world data, which contain times, direction, and movement speed.

\section{Network Prediction}

There are also approaches to predict network traffic.
To forecast network traffic, Ferreira et al.~\cite{ferreiraForecastingNetworkTraffic2023} compare different \ac{ml} models such as \ac{rnn} and \ac{lstm} and perform experiments with real-world data.
Mirza et al.~\cite{mirzaMachineLearningApproach2007} predict \ac{tcp} throughput for arbitrary network paths in the Internet with the \ac{ml} model Support Vector Machines.

\section{Summary}
Various research contributions have been made in wireless communications and networking, focusing on challenges ranging from handover prediction to user mobility and network traffic prediction.
While the abovementioned studies have focused more on sparse \ac{wifi} environments, this thesis focuses on dense \ac{wifi} environments.
To the best of my knowledge, a combination of user movement and handover predictions for large \ac{wifi} environments has not been explored yet, which is addressed in this thesis.
